---
title: "acs"
params: 
  inputDir: "gather/"
  thisDir:  "gather/"
  outputDir: "gathered/"
  fileStem: "acs"
---

```{r KnitrOpts, echo=FALSE}
knitr::opts_knit$set(root.dir = normalizePath('../'))
knitr::opts_chunk$set(echo=TRUE, message=TRUE, warning=TRUE, error=TRUE)
```

```{r}
library(dplyr)
library(acs)
library(tibble)
library(tidyr)
```

To extract this data, you'll need an API key. 

## Geo

We'll be looking at data at the county level for each state. This will be 3,220 rows.

```{r}
geo <- geo.make(state = '*', county="*")
```

```{r}
convert_to_df <- function(obj){
  df <- data.frame(estimate(obj)) %>% 
    rownames_to_column("county_state") %>% 
    separate("county_state", into = c("county", "state"), sep=",")
  df
}
```

## Demographics of interest

First, we'll try to identify some interesting columns. We're going to search for these broad categories:

* Population
* Age
* Education
* Income
* Unemployment
* Size and composition of household

## Population

```{r}
pop_cols <- acs.lookup(keyword="pop", endyear=2015)
dfPopulationCols <- education_cols@results %>% 
  dplyr::filter(grepl("(^|\\W)[Pp]opulation($|\\W)", variable.name))
```

```{r}
obj_pop <- acs.fetch(endyear = 2015, span = 5, geography = geo,
                    table.number = "B16008", col.names = "pretty")
dfPop <- obj_pop %>% 
  convert_to_df()
```

## Age

```{r}
age_cols <- acs.lookup(keyword="age", endyear=2015)
dfAgeCols <- age_cols@results %>% 
  dplyr::filter(grepl("(^|\\W)age($|\\W)", variable.name))
```

```{r eval=FALSE}
View(dfAgeCols)
```

```{r results = 'asis'}
pander::pander(dfAgeCols %>% head())
```

We settle on `B01002`, which has median age by gender and in total and `B01002A` which has median age by white alone.

```{r}
obj_age <- acs.fetch(endyear = 2015, span = 5, geography = geo,
                    table.number = "B01002", col.names = "pretty")
dfAge1 <- obj_age %>% 
  convert_to_df()

obj_age2 <- acs.fetch(endyear = 2015, span = 5, geography = geo,
                    table.number = "B01002A", col.names = "pretty")
dfAge2 <- obj_age2 %>% 
  convert_to_df()
names(dfAge2) <- c("county", "state", paste0(setdiff(names(dfAge2), c("county", "state")), ".White.Alone"))

dfAge <- dplyr::full_join(dfAge1, dfAge2, by = c("county", "state"))
```

## Education

```{r}
education_cols <- acs.lookup(keyword="pop", endyear=2015)
dfEducationCols <- education_cols@results %>% 
  dplyr::filter(grepl("(^|\\W)education($|\\W)", variable.name))
```

```{r eval=FALSE}
View(dfEducationCols)
```

```{r results = 'asis'}
pander::pander(dfEducationCols %>% head())
```

```{r}
geo <- geo.make(state = '*', county="*")
obj_age <- acs.fetch(endyear = 2015, span = 5, geography = geo,
                    table.number = "B01002", col.names = "pretty")
dfAge1 <- data.frame(estimate(obj_age)) %>% 
  rownames_to_column("county_state") %>% 
  separate("county_state", into = c("county", "state"), sep=",")

obj_age2 <- acs.fetch(endyear = 2015, span = 5, geography = geo,
                    table.number = "B01002A", col.names = "pretty")
dfAge2 <- data.frame(estimate(obj_age)) %>% 
  rownames_to_column("county_state") %>% 
  separate("county_state", into = c("county", "state"), sep=",")
names(dfAge2) <- c("county", "state", paste0(setdiff(names(dfAge2), c("county", "state")), ".White.Alone"))

dfAge <- dplyr::full_join(dfAge1, dfAge2, by = c("county", "state"))
```

## 

```{r ListObjects}
lstObjects <- represtools::ListObjects()
represtools::DescribeObjects(represtools::NamesToObjects(lstObjects))
```

```{r Save}
dfAge %>% write.csv(paste0(params$outputDir, "age.csv"), row.names = FALSE)
dfPop %>% write.csv(paste0(params$outputDir, "population.csv"), row.names = FALSE)
save(file = represtools::OutputFile(params)
     , list = lstObjects)
```
